---
title: "Visualization"
author: "Margaret Gacheru"
date: "September 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set up the libraries

```{r}
library(tidyverse)
library(ggridges)
```

Data importing

```{r data_import, cache=TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df


```

Start a plot

```{r}

#aes tells ggplot the specific mapping 
#use the plus sign to tell ggplot what kind of plot you want

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

```

The code below could be used instead to produce the same figure. Using this style can be helpful if you want to do some pre-processing before making your plot but don’t want to save the intermediate data.

```{r}
weather_df %>%
  filter(name == "CentralPark_NY")%>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Adding color

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

Adding a smooth curve (adding a geom) and make the data more transparent

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)

#se = False tells ggplot to remove the standard error bars
#alpha = the level in which a point is transparent [0,1]
```

Have a smooth curve for each group (instead of a global one)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) 
```

Splitting the groups 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

#facet grid splits the groups
```

Size of points based on a specific variable

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

#geom_line connects the dots with a line
```

Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth())

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(tmax_fahr = tmax * (9 / 5) + 32,
         tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)

```

Understanding the distribution of the data (univariate plots)

```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

You can play around with things like the bin width and set the fill color using an aesthetic mapping.

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

The position = "dodge" places the bars for each group side-by-side, but this gets sort of hard to understand.  Density plots in place of histograms might be a better idea

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

Boxplots

```{r}
ggplot(weather_df, aes(y = tmax)) + geom_boxplot()
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot() #separate boxplots by groups


```

Ridge plot

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

Save and embed a plot

```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```



